Rate Transfers with Polyhedral Surrogate Losses /\\ Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-21 rates-neurips-21/rates.tex

Naturally, a central question in surrogate risk minimization is how to choose the surrogate $L$.
A minimal requirement is statistical \emph{consistency}, which roughly says that approaching the best possible surrogate loss will eventually lead to the best possible target loss.
More generally, we would like to understand the rate at which this minimization occurs, called the \emph{generalization rate}: in terms of the number of data points, how quickly the target loss approaches the best possible.
In other words, if we define the \emph{regret} of a hypothesis to be its performance relative to the Bayes optimal hypothesis, we wish to know how quickly the $\ell$-regret approaches zero.

@@

We take a large step toward general surrogate regret bounds, with two broad results.
The first gives a linear surrogate regret bound for any polyhedral (piecewise-linear convex) surrogate, meaning that surrogate generalization rates translate directly to target rates.
  % \begin{theorem}[Linear bounds for polyhedral surrogates]
\begin{theorem}[Linear bound for polyhedral]
  \label{thm:main-upper}
  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
  Then there is a linear regret transfer function $\zeta$.
  % Then $(L,\psi)$ guarantee a linear regret transfer function $\zeta$ to $\ell$.
  %Then there exists a surrogate regret bound with a linear transfer function $\zeta$.
\end{theorem}

@@

\section{Introduction}
\paragraph{The surrogate risk framework.}
In supervised learning, data is of the form $(x,y)$ and the goal is to learn a hypothesis $g(x)$.
This paper considers the common setting where the space of labels $y$ is finite, as is the space of predictions $g(x)$, and accuracy is measured with a \emph{target loss function} $\ell(g(x),y)$.

Because these discrete settings are generally computationally challenging, a common paradigm is to learn $g$ via \emph{surrogate risk minimization}:
\begin{enumerate}
  \item Define a \emph{surrogate loss function} $L(h(x),y)$ over a more-tractable continuous, convex prediction space.
  \item Learn a hypothesis $h(x)$ mapping $x$ to this new prediction space.
  \item Produce a hypothesis $g(x)$ for the original, target problem using some link function $\psi$ via $g(x) = \psi(h(x))$.
\end{enumerate}
For example, in binary classification, the target problem is given by $0$-$1$ loss, $\ell(r)_y = \ones\{r\neq y\}$, with surrogate losses typically defined over $\reals$.
Common surrogate losses include hinge loss and logistic loss.
These surrogates are typically associated with the link $\psi(u) = \mathrm{sign}(u)$, which maps negative predictions to $-1$ and nonnegative predictions to $+1$.

\paragraph{Rates of convergence.}
Given $n$ data points and a surrogate loss $L$, a learning algorithm should produce a hypothesis $h_n$ whose loss converges to the optimal possible as $n$ increases.
Specifically, we can define $\Reg_{L}(h)$ as the expected loss of $h$ on the data distribution, minus the expected loss of the Bayes optimal predictor, i.e. the measurable function with least loss.
Then a primary goal is to bound the \emph{convergence rate} of this algorithm, i.e. provide a bound of the form $\Reg_{L}(h_n) \leq O\left(\tfrac{1}{n^{\alpha}}\right)$.
Such bounds for $\alpha = \tfrac{1}{2}$ are considered ``slow'' rates and are guaranteed by e.g. VC-dimension bounds.\bo{cite}
Faster rates such as $\alpha = 1$ are obtainable is some settings.\bo{cite}

\paragraph{Problem: relating surrogate and target regret.}
However, results of the above form do not guarantee the success of the learning framework for the original, target problem.
Even if the surrogate regret $\Reg_L(h_n)$ is small, we actually care about the induced target hypothesis $g_n = \psi \circ h_n$.
Will the target regret necessarily be small?
Motivated by this problem, this paper will study \emph{surrogate regret bounds}.\bo{cite}
%A surrogate regret bound answers the following question: when we minimize the \emph{surrogate} regret at a given rate, how fast does the \emph{target} regret diminish?
Informally, a surrogate regret bound takes the form
\begin{equation}
  \label{eq:surrogate-regret-bound-informal}
  \forall \text{ hypotheses } h:\X\to\reals^d, \quad \Reg_\ell(\psi\circ h) \leq \zeta(\, \Reg_L(h) \,)~,
\end{equation}
where the regret transfer function $\zeta : \reals_+ \to \reals_+$ controls how surrogate regret translates to target regret.
For example, \bo{cite} studies the \bo{??} surrogate loss for the \bo{??} problem, achieving a guarantee of $\zeta(\epsilon) \leq O\left(\sqrt{\epsilon}\right)$.

While such a guarantee is certainly nice, it is not ideal: even if one achieves a fast rate of $O(\tfrac{1}{n})$ for the surrogate problem, this only translates to the slow rate guarantee of $O(\tfrac{1}{n^{1/2}})$ in the target problem that we actually want to solve.
Preferable would be a transfer function such as $\zeta(\epsilon) = O(\epsilon^{3/4})$ or even $\zeta(\epsilon) = O(\epsilon)$, a linear transfer rate.

\raf{Of particular interest is the concavity of $\zeta$} \bo{explain why?}

By considering a specific class of surrogate losses -- \emph{polyhedral}, generalizations of piecwise-linear convex -- we will show that this utopia is in fact achieved for every surrogate in this class.

%\begin{theorem}[Linear bounds for polyhedral surrogates]
\begin{theorem}[Main surrogate regret bound]
  \label{thm:main-upper}
  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
  Then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$.
  %Then there exists a surrogate regret bound with a linear transfer function $\zeta$.
\end{theorem}

We will also show a class of counterexamples for surrogate losses that are far from polyhedral, i.e. those that are smooth and strongly convex.
This could be viewed as a partial converse, providing evidence that some element of ``piecewise-linearity'' is \emph{necessary} for optimal surrogate regret bounds.
\begin{theorem}[Main lower bound]
  \label{thm:main-lower}
  Let $L$ be a differentiable convex surrogate, with locally Lipschitz gradient, which is locally strongly convex.
  If $L$ satisfies a regret transfer rate of $\zeta$ for a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
\end{theorem}

\bo{example/discussion?}

@@

\bo{other?}

>>>>

Naturally, a central question in surrogate risk minimization is how to choose the surrogate $L$.
A minimal requirement is statistical \emph{consistency}, which roughly says that approaching the best possible surrogate loss will eventually lead to the best possible target loss.
A more precise goal is to understand the rate at which this minimization occurs, called the \emph{generalization rate}: in terms of the number of data points, how quickly the target loss approaches the best possible.
In other words, if we define the \emph{regret} of a hypothesis to be its performance relative to the Bayes optimal hypothesis, we wish to know how quickly the $\ell$-regret approaches zero.

@@

\bo{I don't love the below phrasing. I think I prefer the approach I took of saying, wouldn't it be cool if we could get linear regret bounds? Actually, we can for this special class of surrogates. I'm not sold that this result is a step toward more general bounds. And it's interesting for its own reasons, not just as a step.}
We take a large step toward general surrogate regret bounds, with two broad results.
The first gives a linear surrogate regret bound for any polyhedral (piecewise-linear convex) surrogate, meaning that surrogate generalization rates translate directly to target rates.
  % \begin{theorem}[Linear bounds for polyhedral surrogates]
\begin{theorem}[Linear bound for polyhedral]
  \label{thm:main-upper}
  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
  Then there is a linear regret transfer function $\zeta$.
  % Then $(L,\psi)$ guarantee a linear regret transfer function $\zeta$ to $\ell$.
  %Then there exists a surrogate regret bound with a linear transfer function $\zeta$.
\end{theorem}

@@

@@

%\section{Introduction}
%\paragraph{The surrogate risk framework.}
%In supervised learning, data is of the form $(x,y)$ and the goal is to learn a hypothesis $g(x)$.
%This paper considers the common setting where the space of labels $y$ is finite, as is the space of predictions $g(x)$, and accuracy is measured with a \emph{target loss function} $\ell(g(x),y)$.
%
%Because these discrete settings are generally computationally challenging, a common paradigm is to learn $g$ via \emph{surrogate risk minimization}:
%\begin{enumerate}
%  \item Define a \emph{surrogate loss function} $L(h(x),y)$ over a more-tractable continuous, convex prediction space.
%  \item Learn a hypothesis $h(x)$ mapping $x$ to this new prediction space.
%  \item Produce a hypothesis $g(x)$ for the original, target problem using some link function $\psi$ via $g(x) = \psi(h(x))$.
%\end{enumerate}
%For example, in binary classification, the target problem is given by $0$-$1$ loss, $\ell(r)_y = \ones\{r\neq y\}$, with surrogate losses typically defined over $\reals$.
%Common surrogate losses include hinge loss and logistic loss.
%These surrogates are typically associated with the link $\psi(u) = \mathrm{sign}(u)$, which maps negative predictions to $-1$ and nonnegative predictions to $+1$.
%
%\paragraph{Rates of convergence.}
%Given $n$ data points and a surrogate loss $L$, a learning algorithm should produce a hypothesis $h_n$ whose loss converges to the optimal possible as $n$ increases.
%Specifically, we can define $\Reg_{L}(h)$ as the expected loss of $h$ on the data distribution, minus the expected loss of the Bayes optimal predictor, i.e. the measurable function with least loss.
%Then a primary goal is to bound the \emph{convergence rate} of this algorithm, i.e. provide a bound of the form $\Reg_{L}(h_n) \leq O\left(\tfrac{1}{n^{\alpha}}\right)$.
%Such bounds for $\alpha = \tfrac{1}{2}$ are considered ``slow'' rates and are guaranteed by e.g. VC-dimension bounds.\bo{cite}
%Faster rates such as $\alpha = 1$ are obtainable is some settings.\bo{cite}
%
%\paragraph{Problem: relating surrogate and target regret.}
%However, results of the above form do not guarantee the success of the learning framework for the original, target problem.
%Even if the surrogate regret $\Reg_L(h_n)$ is small, we actually care about the induced target hypothesis $g_n = \psi \circ h_n$.
%Will the target regret necessarily be small?
%Motivated by this problem, this paper will study \emph{surrogate regret bounds}.\bo{cite}
%%A surrogate regret bound answers the following question: when we minimize the \emph{surrogate} regret at a given rate, how fast does the \emph{target} regret diminish?
%Informally, a surrogate regret bound takes the form
%\begin{equation}
%  \label{eq:surrogate-regret-bound-informal}
%  \forall \text{ hypotheses } h:\X\to\reals^d, \quad \Reg_\ell(\psi\circ h) \leq \zeta(\, \Reg_L(h) \,)~,
%\end{equation}
%where the regret transfer function $\zeta : \reals_+ \to \reals_+$ controls how surrogate regret translates to target regret.
%For example, \bo{cite} studies the \bo{??} surrogate loss for the \bo{??} problem, achieving a guarantee of $\zeta(\epsilon) \leq O\left(\sqrt{\epsilon}\right)$.
%
%While such a guarantee is certainly nice, it is not ideal: even if one achieves a fast rate of $O(\tfrac{1}{n})$ for the surrogate problem, this only translates to the slow rate guarantee of $O(\tfrac{1}{n^{1/2}})$ in the target problem that we actually want to solve.
%Preferable would be a transfer function such as $\zeta(\epsilon) = O(\epsilon^{3/4})$ or even $\zeta(\epsilon) = O(\epsilon)$, a linear transfer rate.
%
%\raf{Of particular interest is the concavity of $\zeta$} \bo{explain why?}
%
%By considering a specific class of surrogate losses -- \emph{polyhedral}, generalizations of piecwise-linear convex -- we will show that this utopia is in fact achieved for every surrogate in this class.
%
%%\begin{theorem}[Linear bounds for polyhedral surrogates]
%\begin{theorem}[Main surrogate regret bound]
%  \label{thm:main-upper}
%  Let $L$ be any polyhedral surrogate and $\psi$ any link function such that $(L,\psi)$ is consistent for the target loss $\ell$.
%  Then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$.
%  %Then there exists a surrogate regret bound with a linear transfer function $\zeta$.
%\end{theorem}
%
%We will also show a class of counterexamples for surrogate losses that are far from polyhedral, i.e. those that are smooth and strongly convex.
%This could be viewed as a partial converse, providing evidence that some element of ``piecewise-linearity'' is \emph{necessary} for optimal surrogate regret bounds.
%\begin{theorem}[Main lower bound]
%  \label{thm:main-lower}
%  Let $L$ be a differentiable convex surrogate, with locally Lipschitz gradient, which is locally strongly convex.
%  If $L$ satisfies a regret transfer rate of $\zeta$ for a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%%  If $\zeta$ is a transfer function in a surrogate regret bound for $L$ to a target loss $\ell$, then there exists $c>0$ such that, for arbitrarily small $\epsilon>0$, we have $\zeta(\epsilon) \geq c\cdot\sqrt{\epsilon}$.
%\end{theorem}
%
%\bo{example/discussion?}
